Synthesizing Verified Components for Cyber Assured Systems Engineering \thanks{DISTRIBUTION STATEMENT A. Approved for public release.}

### 2023-02-23 paper/paper.tex
Fix overful boxes

\begin{abstract}
Avionics need to be engineered to be cyber-resilient meaning that systems are able to detect and recover from attacks or safely shutdown.
%
As there are few development tools for cyber-resiliency, designers rely on guidelines and checklists, sometimes missing vulnerabilities until late in the process where remediation is expensive.
%
Our solution is a model-based approach with cyber-resilience-improving transforms that insert high-assurance components such as filters to block malicious data or monitors to detect and alarm anomalous behavior.
%
Novel is our use of model checking and a verified compiler to specify, verify, and synthesize these components.
%
We define \emph{code contracts} as formal specifications that designers write for high-assurance components, and \emph{test contracts} as tests to validate their behavior.
%
A model checker proves whether or not code contracts satisfy test contracts in an iterative development cycle.
%
The same model checker also proves whether or not a system with the inserted components, assuming they adhere to their code contracts, provide the desired cyber-resiliency for the system.
%
We define an algorithm to synthesize implementations for code contracts in a semantics-preserving way that is backed by a verified compiler.
%
The entire workflow is implemented as part of the open source \emph{\brfcs} toolkit.
%
We report on our experience using \brfcs\ with a case study on a UAV system that is transformed to be cyber-resilient to communication and supply chain cyber attacks.
%
Our case study demonstrates that writing code contracts and then synthesizing correct implementations from them is feasible in real-world systems engineering for cyber-resilience.
\end{abstract}

@@

\section{Introduction} \label{sec:intro}
In recent years, aerospace stakeholders have realized that avionics systems are subject to possible cyber-attacks just like other cyber-physical systems.
%
\footnote{This work was funded in part by the
Defense Advanced Research Projects Agency (DARPA) CASE program.
%
The views expressed are those of the authors and do not reflect the official
policy or position of DARPA or the U.S. Government.}
%
Thus, in addition to being fault-tolerant, safety-critical avionics systems must also be {\em cyber-resilient}.
%
Cyber-resiliency means that the system is tolerant to cyber-attacks just as safety-critical systems are tolerant to random faults: they recover and continue to execute their mission function, or safely shut down, as requirements dictate.

Unfortunately, systems designers are currently given few development tools to help answer even basic questions about potential vulnerabilities and ways to mitigate vulnerabilities.
%
They instead rely on process-oriented checklists and guidelines.
%
Cyber vulnerabilities are often discovered during penetration testing late in the development process;
%
or worse yet, they may be discovered only after the product has been fielded, necessitating extremely expensive and time-consuming remediation. This is not a sustainable development model.

@@

The main organizing concept is that of an architecture-to-architecture \emph{security-improving transform}, achieved via the insertion of new architectural components aimed at mitigating cyber-vulnerabilities.
%
We describe two such transformations in this paper: (1) the insertion of a filter to prevent malformed data from a malicious actor being propagated to downstream components, and (2) the insertion of a runtime monitor to detect (and alert) unexpected behaviors arising from uncontrolled components.
%
We refer to these as \emph{high-assurance components} because they must be correct and fulfill their intended purpose for the system to be cyber-resilient.

The contributions of this paper relate to the aspects of \brfcs\ concerning evidence that such high-assurance components meet their intended purpose and are properly implemented.
%
We leverage the existing {\agr} model checker to provide evidence of the former and the existing verified \ckml\ compiler to provide evidence of the latter \cite{agree2013,compositional-analysis-agree,nfm:agree,cakeml}.
%
\footnote{\agr\ is the acronym for \emph{Assume Guarantee Reasoning Environment}.}
%
Specifically, designers write \emph{contracts} to formally specify input and output properties of the system and its components, and \agr\ proves via model checking whether or not the system composition preserves contract requirements at component inputs and guarantees contract obligations at system outputs.
%
In this way, \agr\ provides the formal framework we use to describe the behavior of an inserted high-assurance component, such as a filter or monitor, and then prove if adding such a component to the system improves cyber-resiliency.

@@

We report on a case study applying these transformations with \brfcs\ to an Unmanned Aerial Vehicle (UAV) system that uses the Air Force Research Laboratory's OpenUxAS services for route planning.
%
OpenUxAS, as an open source product, is considered \emph{uncontrolled}.
%
The UAV system is thus transformed to be resilient to malicious behavior that may arise in the uncontrolled component.
%
In particular, filters are added to guard against malformed input and a monitor is created to guard against malicious flight plans generated from OpenUxAS.
%
The case study system is complex and shows the viability of the approach in potential full-scale industrial design.

@@

Our approach currently applies to common cyber-vulnerabilities, such as buffer overflow, lack of input validation, and supply-chain issues.
%
Other cyber-vulnerabilities such as side-channel attacks and denial of service have not yet been addressed. Moreover, this work does not introduce a new type of high-assurance component in terms of capability.
%
Instead, our contribution is in the automated synthesis of security-improving components from formal cyber-resilience specifications, paired with a verification showing that the synthesized implementation is correct in the context of the whole system all the way down to the binary implementation.

@@

\begin{figure*}[h]
  \begin{center}
    \begin{tabular}{c}
      \includegraphics[width=\textwidth]{example.png}
    \end{tabular}
  \end{center}
\caption{Initial design for an automated UAV route planning system.}
\label{fig:example}
\end{figure*}

@@

The expected behavior of the SW system, and the components implementing the system, are modeled with \agr\ contract specifications as shown by the unshaded boxes in the top-left of \figref{fig:workflow}.
%
These contracts assume and guarantee the absence of any malicious, or unspecified, component behavior.
%
More precisely, the contract for SW assumes that its inputs are \emph{well-formed} and that there is never more than one automation request pending at a time.
%
Well-formed generally refers to a syntactic restriction on data at a port. For example, a waypoint is well-formed if it falls within bounds for latitude, longitude, and altitude.
%
The guarantees for SW ensure that
\begin{compactitem}
\item a start coincides with a new waypoint being output;
\item a start is within one cycle of an automation request and if not, then it persistently alerts;
\item new waypoints coincide with location updates; and
\item all outputs are well-formed.
\end{compactitem}

The contracts for the sub-components assume their inputs are well-formed, and they guarantee their outputs are well-formed.
%
The AI contract guarantees it only responds to automation requests and always in the same cycle.
%
The legacy WM contract guarantees that
\begin{compactitem}
  \item it generates a start from a response from the AI and always in the same cycle;
  \item the start coincides with a waypoint; and
  \item any further waypoints coincide with location updates.
\end{compactitem}

This behavioral contract definition of the UAV software system is the starting point on the left-side of \figref{fig:workflow}.
%
The \agr\ verification engine tries to prove if the component composition of the software system is correct.
%
It is an iterative process where the designer updates contracts in the composition until \agr\ proves out the composition.
%
For our example, the initial specifications described above pass verification, meaning that the contract composition of the components with the system satisfies all component input assumptions and system output guarantees.
%
The \agr\ verification conditions, syntax and semantics, and the baseline contracts for the example are discussed in detail in \secref{sec:agree}.
%
\figref{fig:sw} in that section is the \agr\ contract for SW.

@@

Having passed \agr\ verification, the workflow in \figref{fig:workflow} proceeds to the thread analysis as the UAV must actively protect itself from malicious actors.
%
Threat analysis identifies the potential of a supply chain attack through the AI route planner since its third-party.
%
To mitigate the identified supply chain attack, the designer adds a requirement that the system must guard against malicious AI behavior.
%
This requirement is reflected in the component contracts by removing all output guarantees from the AI contract since that component is considered uncontrolled.
%
Without output guarantees, the AI component is able to output arbitrary events in the \agr\ verification analysis.

@@

As \agr\ proves the system now meets its cyber-requirements, the workflow moves to the right in \figref{fig:workflow} to the code generation box.
%
High-assurance components are automatically synthesized by \splt\ from the code contracts to equivalent programs in the \ckml\ language.
%
In other words, for any set of input streams that meet the component's contract assumptions, the output streams produced from the synthesized \ckml\ code exactly match those defined by the contract guarantees.

@@

A system implementation is described by its ports, its subcomponents, and the connections between ports on different subcomponents.
%
Ports in \agr\ are inherited from AADL.
%
There are three kinds of port: \emph{event} ports, \emph{data} ports, and \emph{event data} ports.
%
An event port carries no data and only signals occurrence (or not) of the event.
%
A data port is a sampled port: the data can be read at anytime.
%
An event data port indicates when data is present with an event.

Data flows through the subcomponents of a system in dependency order, with inputs being propagated to outputs through all contracts until they stabilize (\ie, they cannot propagate further).
%
Therefore, subcomponent contracts, and thus the top-level model, must be acyclic.
%
Cyclic systems are made acyclic by breaking cycles with delays.
%
Once the data propagation has stabilized, the model proceeds to the next input data in the input streams.
%
It is worth noting that the semantics do not model computation or communication delay: the output of one contract is seen at the input of any downstream contract in the same step of the input data stream.

@@

A \emph{system} is a tuple $(\inputs, \outputs, \assumes, \guarantees,C)$, where $\inputs$ is the input set, $\outputs$ is the output set, $\assumes$ is the set of assumptions, $\guarantees$ is the set of guarantees, and $C$ are subcomponents.
%
Assumptions and guarantees only use the past-time operators from PLTL.
%
A subcomponent $\component$ is, hierarchically, also a system, and may be designated by its own tuple $(\inputs_\component, \outputs_\component, \assumes_\component, \guarantees_\component, C_c)$.
%
From the components and their connections, $\mathbb{I}_\component$ is defined to be the set of components providing input to some component $\component$ in the system, and $\mathbb{O}$ is defined to be the set of components that provide the output for the system.

@@

The semantics of terms and formulas is defined over \emph{streams of values}.
%
Values encompass at least booleans and numbers but can be readily extended to include records and arrays.
%
A value stream is a total function from time (natural numbers including 0) to values:
%
\[
 \konst{stream} = \mathbb{N}_0 \to \konst{value}
\]
%
Given an \emph{environment} $E : \konst{name} \mapsto \konst{stream}$ binding variable names to value streams, the semantics $\sem{-}^E_t$ of terms and formulas defines the meaning of compound syntax in terms of the meaning of subexpressions.
%
The value of a variable $v$ at time $t$ is found by looking up the stream bound to $v$ in $E$ (call it $s$) and returning $s_t$. Some clauses of the semantics follow, omitting the temporal operators:
%
\[
\begin{array}{rcl}
\sem{v}^E_t & = & E(v)(t) \\
\sem{c}^E_t & = & c \\
\sem{e_1 + e_2}^E_t & = & \sem{e_1}^E_t + \sem{e_2}^E_t \\
   & \cdots & \\
\sem{b_1 \land b_2}^E_t & = & \sem{b_1}^E_t \land \sem{b_2}^E_t \\
   & \cdots & \\
\end{array}
\]

@@

\item
\lineref{line:sw-guarantee-5}:
%
the alert is signaled if the start does not arrive within the one-step bound.
%
Together with \lineref{line:sw-guarantee-4} this allows for non-alerting and alerting behavior.

@@

How does this notion of correctness integrate with the verification conditions generated by \agr?
%
In \agr, the model checker is proving the following property
%
\[
\konst{Hist}(A) \imp P
\]
%
and in \defref{def:correct} we essentially prove a Hoare triple of the form
%
\[
\set{\konst{Hist}(A)}\; (\mathit{Eqs} \cdot G) \; \set{P}
\]
%
This accords with intuition, namely that \agr\ is a kind of \emph{program-free program logic};
%
by pulling out a program $(\mathit{Eqs}\cdot G)$ from a code contract we find ourselves in the domain of imperative program verification.
%
For example, at this point one could apply a verification condition generator to help break down the proof obligation.
%
In order to use a code contract's computed result in verification, we have to show it's defined computation implies the system verification condition as stated by \defref{def:correct}.

@@

\newsavebox{\flt}
\begin{lrbox}{\flt}
  \begin{lstlisting}[style=agree,numbers=left] -- start user
    -- start user definitions
    eq policy : bool =
      WELL_FORMED_AUTOMATION_RESPONSE(Input);
    -- end user definitions

@@

As shown in \figref{fig:hardened-certificate}, the cyber-hardened system passes AGREE verification.
%
That means that the high-assurance components protect the WM from the uncontrolled AI generating malformed data, spontaneous responses, or delayed responses.
%
Synthesis must generate faithful implementations of the code contracts for the \agr\ verification result to apply to the deployed system.

@@

Test contracts are effective in detecting vacuity, under-specified behavior, and missed corner cases.
%
They enable designers to iteratively develop code contracts with \agr\ in the modeling environment, and they build assurance that the code contract computes the desired behavior with black-box, white-box, and other test coverage metrics.
%
These same tests can be synthesized to the backend implementation for traditional testing, if for example, the code contract is a model of an existing legacy binary.

@@

\paragraph{Regular expressions:\/} often the well-formed property of data on connections can be enforced by matching against a regular expression.
%
In other words, the data can be characterized by a \emph{regular language}.
%
As is well known, regular expressions can be translated to efficient deterministic finite state machine (DFA) implementations.
%
In \cite{formal-filter-synth-langsec,case-verified-filter} we report on the creation and application of table-based DFA filters from a verified regular expression-to-DFA compiler.
%
The latter paper offers a fully-worked out example synthesizing a regular expression filter and formally proving that it meets an non-terminating behavioral specification.

@@

When full computational power is needed, we can use the expressive power of \agr\ code contracts to write arbitrarily complex filter or monitor operations, as we have already seen.
%
Our formal model of \agr\ syntax and semantics is applied to extract code contracts and map them to equivalent logic functions.
%
In particular, given a code contract with functionality ${\Eqs}\cdot G$, input ports $\mathbb{I}$, state variables $\mathbb{S}$, and output ports $\mathbb{O}$, the semantics given in \secref{sec:code-contracts} can be applied to define a stream-oriented \emph{step} function $\konst{strmStep} : \mathbb{N}_0 \to \mathit{Env} \to \mathit{Env}$ defined as
%
\[
 \konst{strmStep}\;t \; E = \sem{\mathit{Eqs} \cdot G}^E_t
\]
%
Now suppose ${\Eqs}\cdot G$ is sequentially imperative.
%
Then we can direcctly create a logic function
%
\[
\konst{stateStep} : \mathit{input} \times \mathit{state} \to \mathit{state} \times \mathit{output}
\]
%
corresponding to \konst{strmStep} that takes input and state as parameters and returns an updated state and output.
%
Thus the environment $\mathit{Env}$ holding all previous values of each variable can be replaced by \emph{parameters}, each of which is able to hold only one data element.
%
Assuming $\mathit{E}' = \konst{strmStep}\;t\; \mathit{E}$, we can formally prove that
%
\[
\vdash (\mathit{E}' \vert_{\mathbb{S}},
 \mathit{E}' \vert_{\mathbb{O}}) =
\konst{stateStep} (\mathit{E}\vert_{\mathbb{I}(t)},
                   \mathit{E}\vert_{\mathbb{S}(t)})
\]
%
In other words, projecting out the state and output components from the result of $\konst{strmStep}\;t\; \mathit{E}$ equals the result of applying \konst{stateStep} on the input and state components of $\mathit{E}$ at time $t$.
%
This theorem connects the stream semantics of the code contract to a pure logic function over the values appearing in the stream.
%
From this point we can use tools in the \ckml\ eco-system to jump from the setting of higher order logic to \ckml\ abstract syntax \cite{cakeml-translator}.
%
In this way we have moved from stream semantics to logic function to CakeML evaluation semantics in a property-preserving way.
%
Once in \ckml, further necessary features can be added, namely imperative variables for storing state between component invocations, also invocations of the (verified) foreign function interface of \ckml\ \cite{cakeml-monadic}.
%
Finally, the \ckml\ compiler can be applied to the program.
%
It delivers an executable which can be shown to preserve the original contract.

@@

\subsection{Generating \ckml}
For any of the three specification languages, we take the following general steps to arrive at an executable.
%
\begin{compactenum}
\item Map from specification to computable function in logic, producing a proof of equivalence.
\item Add parsers and printers (specified by contiguity types) for input/output.
\item Map to \ckml\ and add calls to input/output library routines via the \ckml\ foreign function interface.
\item Invoke the \ckml\ compiler.
\end{compactenum}
%
In the following, we examine further details in both filter and monitor synthesis.

@@

The synthesized \ckml\ code for \figref{fig:filter-spec} is in \figref{fig:filter-cakeml}.
%
The code is called at dispatch by the scheduler.
%
The \texttt{API.callFFI} is the link to the communication fabric to capture input and provide output.
%
The body of the function restates the filter contract to make the appropriate assignments in a way that matches the truth value of the predicate in the filter guarantee.

@@

In this section, we outline the application of the \brfcs\ tool towards the development of a UAV surveillance system.
%
The motivation was to showcase the utility and capability of the BriefCASE toolchain on an industrial scale problem.
%
Another part of motivation wast to determine whether or not the technology provides designers with early validation of their models,
%
produces early information about cyber-vulnerabilities,
%
provides effective mitigations,
%
and handles a complex real-world message format, namely the UxAS message format.

@@

We have modeled the system architecture of the UAV in AADL.
%
The model includes a Mission Computer for communicating with the ground station and generating flight plans, and a Flight Control Computer for UAV navigation.
%
The Mission Computer architecture model includes hardware components such as a processor, memory, and communication devices, as well as software.
%
The initial software architecture model (shown in \figref{fig:sw-initial}) contains drivers for communication with the Ground Station and Flight Control Computer, a Waypoint Manager component that provides flight plan coordinates to the Flight Control Computer, and the Flight Planner.
%
The Flight Planner is the open-source UxAS software developed by AFRL~\cite{uxas}.

@@

\begin{figure*}[h]
	\centering
	\includegraphics[width=2\columnwidth]{figs/sw-initial.PNG}
	\caption{Initial software architecture.}
	\label{fig:sw-initial}
\end{figure*}

@@

The intent of the data integrity requirements is to prevent malformed messages from causing a buffer overrun or code injection attack.
%
In the UAV design, such messages are most likely to originate from a remote source or the uncontrolled UxAS component.
%
By placing filters on the connections upstream of mission-critical components, such attacks could be mitigated.
%
We therefore inserted filter components on the incoming and outgoing UxAS connections.

>>>>

\begin{abstract}
Safety critical systems such as avionics need to be engineered to be cyber-resilient meaning that systems are able to detect and recover from attacks or safely shutdown.
%
As there are few development tools for cyber-resiliency, designers rely on guidelines and checklists, sometimes missing vulnerabilities until late in the process where remediation is expensive.
%
Our solution is a model-based approach with cyber-resilience-improving transforms that insert high-assurance components such as filters to block malicious data or monitors to detect and alarm anomalous behavior.
%
Novel is our use of model checking and a verified compiler to specify, verify, and synthesize these components.
%
We define \emph{code contracts} as formal specifications that designers write for high-assurance components, and \emph{test contracts} as tests to validate their behavior.
%
A model checker proves whether or not code contracts satisfy test contracts in an iterative development cycle.
%
The same model checker also proves whether or not a system with the inserted components, assuming they adhere to their code contracts, provide the desired cyber-resiliency for the system.
%
We define an algorithm to synthesize implementations for code contracts in a semantics-preserving way that is backed by a verified compiler.
%
The entire workflow is implemented as part of the open source \emph{\brfcs} toolkit.
%
We report on our experience using \brfcs\ with a case study on a UAV system that is transformed to be cyber-resilient to communication and supply chain cyber attacks.
%
Our case study demonstrates that writing code contracts and then synthesizing correct implementations from them is feasible in real-world systems engineering for cyber-resilience.
\end{abstract}

@@

\section{Introduction} \label{sec:intro}
In recent years, stakeholders have realized that avionics systems are subject to possible cyber-attacks just like other cyber-physical systems.
%
\footnote{This work was funded in part by the
Defense Advanced Research Projects Agency (DARPA) CASE program.
%
The views expressed are those of the authors and do not reflect the official
policy or position of DARPA or the U.S. Government.}
%
Thus, in addition to being fault-tolerant, safety-critical avionics systems must also be {\em cyber-resilient}.
%
Cyber-resiliency means that the system is tolerant to cyber-attacks just as safety-critical systems are tolerant to random faults: they recover and continue to execute their mission function, or safely shut down, as requirements dictate.

Unfortunately, systems designers are given few development tools to help answer even basic questions about potential vulnerabilities and ways to mitigate vulnerabilities.
%
They instead rely on process-oriented checklists and guidelines.
%
Cyber vulnerabilities are often discovered during penetration testing late in the development process;
%
or worse yet, they may be discovered only after the product has been fielded, necessitating extremely expensive and time-consuming remediation. This is not a sustainable development model.

@@

The main concept is that of \emph{security-improving transform} through the insertion of new components aimed at mitigating cyber-vulnerabilities.
%
We describe two such transformations in this paper: (1) the insertion of a filter to prevent malformed data from a malicious actor being propagated to downstream components, and (2) the insertion of a runtime monitor to detect (and alert) unexpected behaviors arising from uncontrolled components.
%
We refer to these as \emph{high-assurance components} because they must be correct and fulfill their intended purpose for the system to be cyber-resilient.

The contributions of this paper relate to the aspects of \brfcs\ concerning evidence that such components meet their intended purpose and are properly implemented.
%
We leverage the existing {\agr} model checker to provide evidence of the former and the existing verified \ckml\ compiler to provide evidence of the latter \cite{agree2013,compositional-analysis-agree,nfm:agree,cakeml}.
%
\footnote{\agr\ is the acronym for \emph{Assume Guarantee Reasoning Environment}.}
%
Specifically, designers write \emph{contracts} to formally specify input and output properties of the system and its components, and \agr\ proves via model checking whether or not the system composition preserves contract requirements at component inputs and guarantees contract obligations at system outputs.
%
In this way, \agr\ provides the formal framework we use to describe the behavior of an inserted high-assurance component, such as a filter or monitor, and then prove if adding such a component to the system improves cyber-resiliency.

@@

We report on a case study applying these transformations with \brfcs\ to an Unmanned Aerial Vehicle (UAV) system.
%
That system uses the Air Force Research Laboratory's OpenUxAS services for route planning.
%
OpenUxAS, as an open source product, is considered \emph{uncontrolled}.
%
The UAV system is thus transformed to be resilient to malicious behavior that may arise in the uncontrolled component.
%
In particular, filters are added to guard against malformed input and a monitor is created to guard against malicious flight plans generated from OpenUxAS.
%
The case study system is complex and shows the viability of the approach in potential full-scale industrial design.

@@

Our approach currently applies to common vulnerabilities such as buffer overflow, lack of input validation, and supply-chain issues.
%
Other cyber-vulnerabilities such as side-channel attacks and denial of service have not yet been addressed. Moreover, this work does not introduce a new type of high-assurance component in terms of capability.
%
Instead, our contribution is in the automated synthesis of security-improving components from formal cyber-resilience specifications, paired with a verification showing that the synthesized implementation is correct in the context of the whole system all the way down to the binary implementation.

@@

\begin{figure*}[ht]
  \begin{center}
    \begin{tabular}{c}
      \includegraphics[width=\textwidth]{example.png}
    \end{tabular}
  \end{center}
\caption{Initial design for an automated UAV route planning system.}
\label{fig:example}
\end{figure*}

@@

The behavior of the SW system, and the components implementing the system, are modeled with \agr\ contract specifications as shown by the unshaded boxes in the top-left of \figref{fig:workflow}.
%
These contracts assume and guarantee the absence of any malicious, or unspecified, component behavior.
%
More precisely, the contract for SW assumes that its inputs are \emph{well-formed} and that there is never more than one automation request pending at a time.
%
Well-formed generally refers to a syntactic restriction on data at a port. For example, a waypoint is well-formed if it falls within bounds for latitude, longitude, and altitude.
%
The guarantees for SW ensure that
\begin{compactitem}
\item a start coincides with a new waypoint being output;
\item a start is within one cycle of an automation request and if not, then it persistently alerts;
\item new waypoints coincide with location updates; and
\item all outputs are well-formed.
\end{compactitem}

The contracts for the sub-components assume their inputs are well-formed, and they guarantee their outputs are well-formed.
%
The AI contract guarantees it only responds to automation requests and always in the same cycle.
%
The legacy WM contract guarantees that
\begin{compactitem}
  \item it generates a start from a response from the AI and always in the same cycle;
  \item the start coincides with a waypoint; and
  \item further waypoints coincide with location updates.
\end{compactitem}

This behavioral definition of the UAV software system is the starting point on the left-side of \figref{fig:workflow}.
%
The \agr\ verification engine tries to prove if the component composition of the software system is correct.
%
It is an iterative process where the designer updates contracts in the composition until \agr\ proves out the composition.
%
For our example, the initial specifications described above pass verification, meaning that the contract composition of the components with the system satisfies all component input assumptions and system output guarantees.
%
The \agr\ verification conditions, syntax and semantics, and the baseline contracts for the example are discussed in detail in \secref{sec:agree}.
%
\figref{fig:sw} in that section is the \agr\ contract for SW.

@@

After \agr\ verification, the workflow in \figref{fig:workflow} proceeds to the thread analysis as the UAV must actively protect itself from malicious actors.
%
Threat analysis identifies the potential of a supply chain attack through the AI route planner since its third-party.
%
To mitigate the identified supply chain attack, the designer adds a requirement that the system must guard against malicious AI behavior.
%
This requirement is reflected in the component contracts by removing all output guarantees from the AI contract since that component is considered uncontrolled.
%
Without output guarantees, the AI component is able to output arbitrary events in the \agr\ verification analysis.

@@

As \agr\ verifies the system now meets its cyber-requirements, the workflow moves to the right in \figref{fig:workflow} to the code generation box.
%
High-assurance components are automatically synthesized by \splt\ from the code contracts to equivalent programs in the \ckml\ language.
%
In other words, for any set of input streams that meet the component's contract assumptions, the output streams produced from the synthesized \ckml\ code exactly match those defined by the contract guarantees.

@@

A system implementation is described by its ports, its sub-components, and the connections between ports on different sub-components.
%
Ports in \agr\ are inherited from AADL.
%
There are three kinds of port: \emph{event} ports, \emph{data} ports, and \emph{event data} ports.
%
An event port carries no data and only signals occurrence (or not) of the event.
%
A data port is a sampled port: the data can be read at anytime.
%
An event data port indicates when data is present with an event.

Data flows through the sub-components of a system in dependency order: the inputs propagate to outputs through the contracts until everything stabilizes.
%
Therefore, sub-component contracts, and thus the top-level model, must be acyclic.
%
Cyclic systems are made acyclic by breaking cycles with delays.
%
Once the data propagation has stabilized, the model proceeds to the next input data in the input streams.
%
It is worth noting that the semantics do not model computation or communication delay: the output of one contract is seen at the input of any downstream contract in the same step of the input data stream.

@@

A \emph{system} is a tuple $(\inputs, \outputs, \assumes, \guarantees,C)$, where $\inputs$ is the input set, $\outputs$ is the output set, $\assumes$ is the set of assumptions, $\guarantees$ is the set of guarantees, and $C$ are sub-components.
%
Assumptions and guarantees only use the past-time operators from PLTL.
%
A sub-component $\component$ is, hierarchically, also a system, and may be designated by its own tuple $(\inputs_\component, \outputs_\component, \assumes_\component, \guarantees_\component, C_c)$.
%
From the components and their connections, $\mathbb{I}_\component$ is defined to be the set of components providing input to some component $\component$ in the system, and $\mathbb{O}$ is defined to be the set of components that provide the output for the system.

@@

The semantics of terms and formulas is defined over \emph{streams of values}.
%
Values encompass at least booleans and numbers but can be readily extended to include records and arrays.
%
A value stream is a total function from time (natural numbers including 0) to values:
%
\[
 \konst{stream} = \mathbb{N}_0 \to \konst{value}
\]
%
Given an \emph{environment} $E : \konst{name} \mapsto \konst{stream}$ binding variable names to value streams, the semantics $\sem{-}^E_t$ of terms and formulas defines the meaning of compound syntax in terms of the meaning of sub-expressions.
%
The value of a variable $v$ at time $t$ is found by looking up the stream bound to $v$ in $E$ (call it $s$) and returning $s_t$. Some clauses of the semantics follow, omitting the temporal operators:
%
\[
\begin{array}{rcl}
\sem{v}^E_t & = & E(v)(t) \\
\sem{c}^E_t & = & c \\
\sem{e_1 + e_2}^E_t & = & \sem{e_1}^E_t + \sem{e_2}^E_t \\
   & \cdots & \\
\sem{b_1 \land b_2}^E_t & = & \sem{b_1}^E_t \land \sem{b_2}^E_t \\
   & \cdots & \\
\end{array}
\]

@@

\item
\lineref{line:sw-guarantee-5}:
%
the alert is signaled if the start does not arrive within one-step.
%
Together with \lineref{line:sw-guarantee-4} this allows for non-alerting and alerting behavior.

@@

How does this definition of code contract correctness integrate with the verification conditions generated by \agr?
%
In \agr, the model checker is proving the following property
%
\[
\konst{Hist}(A) \imp P
\]
%
and in \defref{def:correct} we essentially prove a Hoare triple of the form
%
\[
\set{\konst{Hist}(A)}\; (\mathit{Eqs} \cdot G) \; \set{P}
\]
%
This accords with intuition, namely that \agr\ is a kind of \emph{program-free program logic};
%
by pulling out a program $(\mathit{Eqs}\cdot G)$ from a code contract we find ourselves in the domain of imperative program verification.
%
For example, at this point one could apply a verification condition generator to help break down the proof obligation.
%
In order to use a code contract's computed result in verification, we have to show it's defined computation implies the system verification condition as stated by \defref{def:correct}.

@@

\newsavebox{\flt}
\begin{lrbox}{\flt}
  \begin{lstlisting}[style=agree,numbers=left]
    -- start user definitions
    eq policy : bool =
      WELL_FORMED_AUTOMATION_RESPONSE(Input);
    -- end user definitions

@@

\figref{fig:hardened-certificate} shows the hardened system to pass AGREE verification.
%
That means that the high-assurance components protect the WM from the uncontrolled AI generating malformed data, spontaneous responses, or delayed responses.
%
Synthesis must generate faithful implementations of the code contracts for the \agr\ verification result to apply to the deployed system.

@@

Test contracts are a tool to help detecting vacuity, under-specification, and missed corner cases.
%
They enable designers to iteratively develop code contracts with \agr\ in the modeling environment, and they build assurance that the code contract computes the desired behavior with black-box, white-box, and other test coverage metrics.
%
These same tests can be synthesized to the backend implementation for traditional testing, if for example, the code contract is a model of an existing legacy binary.

@@

\paragraph{Regular expressions:\/} often the well-formed property of data can be enforced by matching against a regular expression.
%
In other words, the data can be characterized by a \emph{regular language}.
%
As is well known, regular expressions can be translated to efficient deterministic finite state machine (DFA) implementations.
%
In \cite{formal-filter-synth-langsec,case-verified-filter} we report on the creation and application of table-based DFA filters from a verified regular expression-to-DFA compiler.
%
The latter paper offers a fully-worked out example synthesizing a regular expression filter and formally proving that it meets an non-terminating behavioral specification.

@@

When full computational power is needed, we can use the expressive power of \agr\ code contracts to write arbitrarily complex filter or monitor operations, as we have already seen.
%
Our formal model of \agr\ syntax and semantics is applied to extract code contracts and map them to equivalent logic functions.
%
In particular, given a code contract with functionality ${\Eqs}\cdot G$, input ports $\mathbb{I}$, state variables $\mathbb{S}$, and output ports $\mathbb{O}$, the semantics given in \secref{sec:code-contracts} can be applied to define a stream-oriented \emph{step} function $\konst{strmStep} : \mathbb{N}_0 \to \mathit{Env} \to \mathit{Env}$ defined as
%
\[
 \konst{strmStep}\;t \; E = \sem{\mathit{Eqs} \cdot G}^E_t
\]
%
Now suppose ${\Eqs}\cdot G$ is sequentially imperative.
%
Then we can directly create a logic function
%
\[
\konst{stateStep} : \mathit{input} \times \mathit{state} \to \mathit{state} \times \mathit{output}
\]
%
corresponding to \konst{strmStep} that takes input and state as parameters and returns an updated state and output.
%
Thus the environment $\mathit{Env}$ holding all previous values of each variable can be replaced by \emph{parameters}, each of which is able to hold only one data element.
%
Assuming $\mathit{E}' = \konst{strmStep}\;t\; \mathit{E}$, we can formally prove that
%
\[
\vdash (\mathit{E}' \vert_{\mathbb{S}},
 \mathit{E}' \vert_{\mathbb{O}}) =
\konst{stateStep} (\mathit{E}\vert_{\mathbb{I}(t)},
                   \mathit{E}\vert_{\mathbb{S}(t)})
\]
%
In other words, projecting out the state and output components from the result of $\konst{strmStep}\;t\; \mathit{E}$ equals the result of applying \konst{stateStep} on the input and state components of $\mathit{E}$ at time $t$.
%
This theorem connects the stream semantics of the code contract to a pure logic function over the values appearing in the stream.
%
From this point we can use tools in the \ckml\ eco-system to jump from the setting of higher order logic to \ckml\ abstract syntax \cite{cakeml-translator}.
%
In this way we have moved from stream semantics to logic function to CakeML evaluation semantics in a property-preserving way.
%
Once in \ckml, further necessary features can be added, namely imperative variables for storing state between component invocations, also invocations of the (verified) foreign function interface of \ckml\ \cite{cakeml-monadic}.
%
Finally, the \ckml\ compiler can be applied to the program.
%
It delivers an executable which can be shown to preserve the original contract.

@@

\subsection{Generating \ckml}
For any of the three specification languages, we take the following general steps to arrive at an executable.
%
\begin{compactenum}
\item Map from specification to computable function in logic, producing a proof of equivalence.
\item Add needed parsers and printers (specified by contiguity types) for input/output.
\item Map to \ckml\ and add calls to input/output library routines via the \ckml\ foreign function interface.
\item Invoke the \ckml\ compiler.
\end{compactenum}
%
In the following, we examine further details in both filter and monitor synthesis.

@@

The synthesized code for \figref{fig:filter-spec} is shown in \figref{fig:filter-cakeml}.
%
The code is called at dispatch by the scheduler.
%
The \texttt{API.callFFI} is the link to the communication fabric to capture input and provide output.
%
The body of the function restates the filter contract to make the appropriate assignments in a way that matches the truth value of the predicate in the filter guarantee.

@@

In this section, we outline the application of the \brfcs\ tool towards the development of a UAV surveillance system.
%
The motivation was to showcase the utility and capability of the BriefCASE toolchain on an industrial scale problem.
%
Another part of motivation wast to determine whether or not the technology provides designers with early validation of their models,
%
produces early information about cyber-vulnerabilities,
%
provides effective mitigation,
%
and handles a complex real-world message format, namely the UxAS message format.

@@

We modeled the system architecture of the UAV in AADL.
%
The model includes a Mission Computer for communicating with the ground station and generating flight plans, and a Flight Control Computer for UAV navigation.
%
The Mission Computer architecture model includes hardware components such as a processor, memory, and communication devices, as well as software.
%
The initial software architecture model (shown in \figref{fig:sw-initial}) contains drivers for communication with the Ground Station and Flight Control Computer, a Waypoint Manager component that provides flight plan coordinates to the Flight Control Computer, and the Flight Planner.
%
The Flight Planner is the open-source UxAS software developed by AFRL~\cite{uxas}.

@@

\begin{figure*}[ht]
	\centering
	\includegraphics[width=2\columnwidth]{figs/sw-initial.PNG}
	\caption{Initial software architecture.}
	\label{fig:sw-initial}
\end{figure*}

@@

The data integrity requirements serve to prevent any malformed messages from causing a buffer overrun or code injection attack.
%
In the UAV design, such messages are most likely to originate from a remote source or the uncontrolled UxAS component.
%
By placing filters on the connections upstream of mission-critical components, such attacks could be mitigated.
%
We therefore inserted filter components on the incoming and outgoing UxAS connections.

The code contracts for these filters are similar in that they check that values contained in the messages are within appropriate ranges.
%
For example, the Automation Response message filter, which drops messages containing malformed flight plans, is defined as shown in \figref{fig:automation-response-filter}.
%
Although \texttt{Latitude}, \texttt{Longitude} and \texttt{Altitude} are defined as 64-bit floating-point values, the filter only passes messages containing waypoint values between [-90,90], [-180,180], and [0,15000], respectively.

\begin{figure}[ht]
	\centering
	\includegraphics[width=1\columnwidth]{figs/automation-response-filter.png}
	\caption{Automation Response Filter specification.}
	\label{fig:automation-response-filter}
\end{figure}
